---
title: "Econ 1042 PS2"
author: "Owen Asnis"
date: "2023-02-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
```

# Data Wrangling 

```{r data wrangling, message=FALSE}

players <- read_delim("nhlps2.csv", delim = "\t", 
    escape_double = FALSE, trim_ws = TRUE)

standings <- read_csv("standings_2018_2019.csv", 
                      show_col_types = FALSE) %>% 
  select(-c(...11, ...12, ...13, ...14, ...15))

```

# Question 1 

### In baseball, expected home runs is a better representation than actual home runs because the expected statistics account for other variables. For example, one batter could be making hard contact but only have a few home runs because he plays in Pittsburgh (ranked by Bleacher Report as the worst ballpark for hitters), whereas another batter could be making softer contact but have more home runs because he plays in Colorado (considered to be the best ballpark for hitters). Even though the player from Pittsburgh has less home runs, we would still consider him to be a better hitter, as he would have more home runs than the player from Colorado if they played in the same conditions. In hockey, the total shots by each team while a player is on the ice might be a better indicator of a player's value than goals and assists because it similarly accounts for other variables. For example, one player could be generating many shots for and allowing few shots against but have scored no goals because his team was playing the best goalie in the league, whereas another player could be generating few shots for and allowing many shots against and scored a goal because his team was playing the worst goalie in the league. Again, the player generating many shots for and few shots against would be considered the better player because he would be more productive than the player generating few shots for and many shots against if they were playing in the same conditions. In basketball, thinking about shots doesn't work in the same way, because after one team scores, the other team gets the ball, leading to very similar shot totals. In hockey, if one team is dominating the other, there will be a large difference in shot totals. 

# Question 2 

### 